---
title: "Data Exploration"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: no
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r echo = F, message = F, warning = F}
library(COVIDYPLL)
library(ggplot2)

dat_ls <- data(package = "COVIDYPLL")
dat_ls <- dat_ls$results[, "Item"]
data(list = dat_ls, package = "COVIDYPLL")
```


# Suppressed and non-suppressed COVID-19 deaths

## Overall

```{r echo = F, message = F, warning = F, fig.width = 8, fig.height = 6}
# Calculate suppress values in each quarter
mort2020[, suppress_covid_19_deaths := ifelse(is.na(covid_19_deaths), "suppressed",
                                             ifelse(covid_19_deaths == 0, "non-suppressed: 0", "non-suppressed: positive"))]
mort2020[, suppress_total_deaths := ifelse(is.na(total_deaths), "suppressed",
                                             ifelse(total_deaths == 0, "non-suppressed: 0", "non-suppressed: positive"))]

sum_suppress <- mort2020[, list(N = .N), by = .(quarter, suppress_covid_19_deaths)]
sum_suppress[, pct := round(N / sum(N) * 100, 1), by = .(quarter)]
sum_suppress[, label := paste0(pct, "%")]
sum_suppress$label[sum_suppress$quarter == 1 & sum_suppress$suppress_covid_19_deaths == "non-suppressed: positive"] <- NA
sum_suppress <- sum_suppress[order(quarter, suppress_covid_19_deaths)]

ggplot(data = sum_suppress, aes(x = quarter, y = pct)) +
  geom_bar(aes(fill = suppress_covid_19_deaths),
           color = "gray30", stat = "identity", position = "stack", size = 0.5) +
  geom_text(aes(x = quarter, y = pct, label = label,
                group = suppress_covid_19_deaths),
            stat = "identity", position = "stack", vjust = 1.5) +
  scale_fill_manual(values = set_colors(3)) +
  ylab("%") +
  ggtitle("Distribution of suppressed and non-suppressed COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        strip.text.x = element_text(size = 14, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.position = "bottom")
```


## State

```{r echo = F, message = F, warning = F, fig.width = 14, fig.height = 10}
mort2020[, suppress_covid_19_deaths := ifelse(is.na(covid_19_deaths), "suppressed",
                                             ifelse(covid_19_deaths == 0, "non-suppressed: 0", "non-suppressed: positive"))]
sum_suppress <- mort2020[, list(N = .N), by = .(state, quarter, suppress_covid_19_deaths)]
sum_suppress[, pct := round(N / sum(N) * 100, 1), by = .(state, quarter)]
sum_suppress[, label := paste0(pct, "%")]
sum_suppress <- sum_suppress[order(quarter, suppress_covid_19_deaths)]

ggplot(data = sum_suppress, aes(x = quarter, y = pct)) +
  geom_bar(aes(fill = suppress_covid_19_deaths),
           color = "gray30", stat = "identity", position = "stack", size = 0.5) +
  scale_fill_manual(values = set_colors(3)) +
  facet_wrap(.~state, scale = "free", ncol = 10) +
  ylab("%") +
  ggtitle("Distribution of suppressed and non-suppressed COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```


## Urban and rural code

```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 6}
mort2020[, suppress_covid_19_deaths := ifelse(is.na(covid_19_deaths), "suppressed",
                                              ifelse(covid_19_deaths == 0, "non-suppressed: 0", "non-suppressed: positive"))]
sum_suppress <- mort2020[, list(N = .N), by = .(urban_rural_code, quarter, suppress_covid_19_deaths)]
sum_suppress[, pct := round(N / sum(N) * 100, 1), by = .(urban_rural_code, quarter)]
sum_suppress[, label := ifelse(pct < 4, NA, paste0(pct, "%"))]
sum_suppress <- sum_suppress[order(quarter, suppress_covid_19_deaths)]

ggplot(data = sum_suppress, aes(x = quarter, y = pct)) +
  geom_bar(aes(fill = suppress_covid_19_deaths),
           color = "gray30", stat = "identity", position = "stack", size = 0.5) +
  geom_text(aes(x = quarter, y = pct, label = label,
                group = suppress_covid_19_deaths),
            stat = "identity", position = "stack", vjust = 1.5) +
  scale_fill_manual(values = set_colors(3)) +
  facet_wrap(.~urban_rural_code, scale = "free") +
  ylab("%") +
  ggtitle("Distribution of suppressed and non-suppressed COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```


# Distribution of COVID-19 deaths

## Log transformed

```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 8}
mort2020[, l_covid_19_deaths := log(covid_19_deaths)]

ggplot(data = mort2020) +
  geom_histogram(aes(x = l_covid_19_deaths, y = ..density..),
                 fill = set_colors(1)) +
  geom_vline(xintercept = log(9), color = "gray20", size = 0.3) +
  scale_x_continuous(breaks = c(0:8), labels = c(0:8)) +
  facet_wrap(.~quarter, scale = "free_y") +
  ggtitle("Distribution of COVID-19 deaths, log transformed") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```


## Original scale <= 200 cases

```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 8}
ggplot(data = mort2020[covid_19_deaths <= 200 & covid_19_deaths > 0]) +
  geom_histogram(aes(x = covid_19_deaths, y = ..density..),
                 fill = set_colors(1)) +
  geom_vline(xintercept = 10, color = "gray20", size = 0.3) +
  scale_x_continuous(breaks = c(0, 10, seq(20, 200, 20)), labels = c(0, 10, seq(20, 200, 20))) +
  facet_wrap(.~quarter, scale = "free_y") +
  ggtitle("Distribution of COVID-19 deaths (<= 200 cases)") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```

# Relationship between COVID-19 deaths, total deaths, and population

```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 8}
ggplot(data = mort2020) +
  geom_point(aes(y = covid_19_deaths, x = total_deaths)) +
  facet_wrap(.~quarter, scale = "free") +
  ggtitle("Relationship between total deaths and COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```

```{r echo = F, message = F, warning = F}
table(mort2020$suppress_total_deaths, mort2020$suppress_covid_19_deaths)
```


```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 8}
ggplot(data = mort2020) +
  geom_point(aes(y = covid_19_deaths, x = pop_size)) +
  facet_wrap(.~quarter, scale = "free") +
  ggtitle("Relationship between population size and COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```


```{r echo = F, message = F, warning = F, fig.width = 10, fig.height = 8}
ggplot(data = mort2020) +
  geom_point(aes(y = total_deaths, x = pop_size)) +
  facet_wrap(.~quarter, scale = "free") +
  ggtitle("Relationship between population size and COVID-19 deaths") +
  theme_bw() +
  theme(plot.title = element_text(size = 18, hjust = 0.5),
        strip.text.x = element_text(size = 12, colour = "gray20"),
        strip.background = element_rect(colour = NA, fill = "white"),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom")
```





